---
title: "QCIRMS Tutorial"
author: "Brett McKinney and Lily Clough"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
  dxf_path: !r "220318_1%_CO2_NaCl_NaSO4_brines"
  plot_path: !r "plots"
  write_pdf: false
always_allow_html: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 4.5,
  fig.align = "center"
)
```

# Overview

This tutorial walks through a practical QCIRMS workflow for continuous-flow isotope ratio
mass spectrometry (IRMS) data stored in Thermo `.dxf` files. It is adapted from the
package pipeline script and reorganized as a reproducible R Markdown analysis.

The workflow has four parts:

1. install and load dependencies;
2. inspect a single `.dxf` file;
3. process a directory of files and generate chromatogram plots;
4. run the `QAQC_IRMS()` directory-level QA/QC workflow.

# Installation

QCIRMS currently depends on `isoreader`, which is installed from GitHub.

```{r install, eval=FALSE}
install.packages(c("dplyr", "pracma"))
install.packages("pak")

pak::pak("isoverse/isoreader")
pak::pak("insilico/QCIRMS")
```

# Load packages

```{r libraries}
library(QCIRMS)
library(isoreader)
library(dplyr)
library(pracma)
```

# Example 1: Inspect one packaged DXF file

QCIRMS ships with a small example DXF file in `inst/extdata`.

```{r single-file-setup}
dxf_file <- "170506_NaHCO3 L + NaCl U_.dxf"
file_path <- system.file("extdata", "dxf_files", "abiotic", dxf_file, package = "QCIRMS")

file_path
```

## Read the file summary

```{r read-summary}
file_summ <- read_summary(file_path)
str(file_summ, max.level = 1)
```

## Vendor peak table

```{r vendor-info}
vend_df <- vendor_info(file_path)
dplyr::glimpse(vend_df)
head(vend_df[, 1:min(10, ncol(vend_df))])
```

## Raw chromatogram data

```{r raw-data}
raw_df <- raw_data(file_path)
dplyr::glimpse(raw_df)
head(raw_df)
```

## Reference values

```{r reference-values}
stand_ratio_df <- reference_values_ratio(file_path)
stand_no_ratio_df <- reference_values_no_ratio(file_path)

head(stand_ratio_df)
head(stand_no_ratio_df)
```

## Plot one chromatogram

```{r one-plot}
generic_raw_plot(
  raw.df = raw_df,
  title = dxf_file
)
```

# Example 2: Batch processing for a directory of DXF files

For a local experiment directory, set `data_dir` to a folder containing `.dxf` files.
If you knit this document without parameters, the chunk below is skipped.

```{r batch-setup}
DATA_PATH <- params$data_dir
RESULTS_PATH <- params$results_dir

if (is.null(RESULTS_PATH) && !is.null(DATA_PATH)) {
  RESULTS_PATH <- file.path(DATA_PATH, "QCresults")
  dir.create(RESULTS_PATH, recursive = TRUE, showWarnings = FALSE)
}
```

```{r batch-processing, eval=!is.null(params$data_dir)}
file_names <- all_dxf_files(path = DATA_PATH)
length(file_names)
head(file_names)

raw_list <- raw_data_all(file_names, path = DATA_PATH)
length(raw_list)
```

```{r batch-plotting, eval=!is.null(params$data_dir)}
generic_plot_all_raw(
  raw.list = raw_list,
  path = RESULTS_PATH,
  write_pdf = TRUE,
  pdf_name = "all_dxf.pdf"
)
```

# Example 3: Directory-level QA/QC with expected reference times

To run the full pipeline you need:

- a directory of `.dxf` files;
- a table of expected reference-peak times, for example
  `referencePeaks_expectedTimes.txt`;
- an output directory for results.

```{r qaqc-inputs}
REF_TIME_FILE <- params$ref_time_file

if (!is.null(REF_TIME_FILE)) {
  expRef_df <- read.table(REF_TIME_FILE, header = TRUE)
  expRef_df
}
```

## Run `QAQC_IRMS()`

```{r run-qaqc, eval=!is.null(params$data_dir) && !is.null(params$ref_time_file)}
start_time <- Sys.time()

curr_filtered <- QAQC_IRMS(
  unfilteredPath = DATA_PATH,
  expRef.df = expRef_df,
  checkIntStand = TRUE,
  internalStandID = c("L1", "H1", "LW"),
  dataName = basename(DATA_PATH),
  maxPkNum = 18,
  expectedNonSampPks = 7,
  sdCrefIso.thresh = 0.1,
  sdOrefIso.thresh = 0.1,
  checkRelDiffIntensity = TRUE,
  refSamp_relDiffInt.thresh = 0.75,
  amplName = "Ampl44",
  ref_relDiffInt.thresh = 0.1,
  sdCsampIso.thresh = 0.3,
  sdOsampIso.thresh = 0.2,
  outPath = RESULTS_PATH,
  verbose = TRUE
)

elapsed <- Sys.time() - start_time
elapsed
```

## Extract QA/QC outputs

```{r qaqc-outputs, eval=!is.null(params$data_dir) && !is.null(params$ref_time_file)}
refs_dat <- curr_filtered[[1]]
samps_dat <- curr_filtered[[2]]

dim(refs_dat)
dim(samps_dat)

head(refs_dat)
head(samps_dat)
```

## Internal standards (optional)

```{r internal-standards, eval=!is.null(params$data_dir) && !is.null(params$ref_time_file)}
int_stand_list <- curr_filtered[[3]]

length(int_stand_list)
str(int_stand_list, max.level = 1)
```

# Suggested next steps

- Add this tutorial as a package vignette under `vignettes/`.
- Replace any hard-coded paths with `params` or `here::here()`.
- Keep only small demonstration files in `inst/extdata/`.
- Store larger local data outside the installed package and document the expected directory structure.

# Session information

```{r session-info}
sessionInfo()
```
