---
title: "Test Report"
author: "QCIRMS"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '3'
params:
  MULT_DXF_DIRS: yes
  DATA_PATH: Users/lily-clough/Documents/data_processing_irms/OW_Schlenk_IsoData/OW_Schlenk_IsoData/OW_Schlenk_IsoData_dxf
  REF_TIME_PATH: ~/Documents/Papers/qcirms/test_220318/220318_1pct_CO2_NaCl_NaSO4_brines/qc
  RESULT_PATH: ~/Documents/Papers/qcirms/test_220318/results_220318
  REF_FILE: referencePeaks_expectedTimes.txt
  CHECK_INTERNAL_STANDARDS: yes
  RESULTS_NAME: 220318_1pct_CO2_NaCl_NaSO4_brines 
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)
library(QCIRMS)
library(isoreader)
library(dplyr)
library(knitr)
```

# Overview

This report was generated automatically from a QCIRMS processing template.
It follows the batch-processing structure of the QCIRMS workflow and includes both written output files and embedded summaries.

# Paths

```{r paths}
dir.create(params$RESULT_PATH, recursive = TRUE, showWarnings = FALSE)
LOOP <- params$MULT_DXF_DIRS
DATA_PATH <- params$DATA_PATH
# LAC check if this is a dir of dxf files or dir of dirs
if(LOOP){
  dxf_dirs <- list.dirs(DATA_PATH,recursive=FALSE)
  #dxf_dirs <- dxf_dirs[!grepl("/\\.",dxf_dirs)]
  print(paste("Found ",as.character(length(dxf_dirs)), " directories of data",sep=""))
  # LAC just grab first for now, figure out how to loop
  # LAC: for loop here, change index to i, loop idex is length(dxf_dirs)
  # re-processes DATA-PATH at the beginning of every loop as above
  DATA_PATH <- dxf_dirs[1]
  #cat("dxf_dirs :", dxf_dirs,"\n")
  #print(DATA_PATH)
  #PLOTS_PATH <- paste(params$RESULT_PATH,"/test_plots",sep="")
  #dir.create(PLOTS_PATH,recursive = TRUE, showWarnings = FALSE)
  #RESULT_PATH <- paste(params$RESULT_PATH,"/test_results",sep="")
  RESULT_PATH <- params$RESULT_PATH
  dir.create(RESULT_PATH,recursive = TRUE, showWarnings = FALSE)
  # grab dxf dir name only
  #RESULTS_NAME <- sub(".*[/\\\\\\\\]", "",DATA_PATH)
  RESULTS_NAME <- params$RESULTS_NAME
  #cat("rn: ", RESULTS_NAME,"\n")
  #dir.create(paste(RESULT_PATH,"/",RESULTS_NAME,sep=""),recursive = TRUE, showWarnings = FALSE)
  dir.create(paste(RESULT_PATH,"/",sep=""),recursive = TRUE, showWarnings = FALSE)

  }

REF_TIME_PATH <- params$REF_TIME_PATH
PLOTS_PATH <- RESULT_PATH#params$PLOTS_PATH
#RESULT_PATH <- paste(RESULT_PATH,"/",RESULTS_NAME,sep="")#params$RESULT_PATH
RESULT_PATH <- paste(RESULT_PATH,"/",sep="")
REF_FILE <- params$REF_FILE
DATA_PATH <- params$DATA_PATH
CHECK_INTERNAL_STANDARDS <- params$CHECK_INTERNAL_STANDARDS
# LAC make output file names dynamic, R double escape for backslashes
#dir.create(RESULT_PATH, recursive = TRUE, showWarnings = FALSE)

path_tbl <- data.frame(
  Setting = c('DATA_PATH', 'REF_TIME_PATH', 'PLOTS_PATH', 'RESULT_PATH', 'REF_FILE', 'CHECK_INTERNAL_STANDARDS'),
  Value = c(DATA_PATH, REF_TIME_PATH, PLOTS_PATH, RESULT_PATH, REF_FILE, CHECK_INTERNAL_STANDARDS),
  stringsAsFactors = FALSE
)
knitr::kable(path_tbl, caption = 'Report parameters')
```

# Discover DXF files

```{r discover-files}
fileNames <- all_dxf_files(path = DATA_PATH)
cat('Number of DXF files found:', length(fileNames), '\n')
if (length(fileNames) > 0) {
  knitr::kable(head(data.frame(file = fileNames, stringsAsFactors = FALSE), 10),
               caption = 'First DXF files discovered')
} else {
  stop('No .dxf files found in DATA_PATH.')
}
```

# Single-file preview

```{r single-file-preview}
first_file <- fileNames[1]
cat('Preview file:', first_file, '\n')

vend.df <- vendor_info(file.path(DATA_PATH, first_file))
raw.df  <- raw_data(file.path(DATA_PATH, first_file))

knitr::kable(head(vend.df[, seq_len(min(10, ncol(vend.df))), drop = FALSE]),
             caption = 'Vendor information preview')
knitr::kable(head(raw.df), caption = 'Raw chromatogram preview')

write.csv(vend.df, file.path(RESULT_PATH, 'single_file_vendor_info.csv'), row.names = FALSE)
write.csv(raw.df,  file.path(RESULT_PATH, 'single_file_raw_data.csv'), row.names = FALSE)
```

```{r single-file-plot, fig.width=8, fig.height=5}
if(dim(raw.df)[1] > 0){ # check for data
  generic_raw_plot(raw.df = raw.df, title = first_file)
}
```

# Reference information

```{r reference-info}
expRef.df <- read.table(file.path(REF_TIME_PATH, REF_FILE), header = TRUE)
knitr::kable(expRef.df, caption = 'Expected reference peak times')
write.csv(expRef.df, file.path(RESULT_PATH, 'expected_reference_times.csv'), row.names = FALSE)
```

# Batch raw-data processing

```{r batch-raw}
rawList <- raw_data_all(fileNames, path = DATA_PATH)
cat('Number of raw chromatogram objects:', length(rawList), '\n')
raw_lengths <- vapply(rawList, function(x) if (is.null(x)) 0L else nrow(x), integer(1))
raw_tbl <- data.frame(file = fileNames, n_rows = raw_lengths, stringsAsFactors = FALSE)
knitr::kable(head(raw_tbl, 10), caption = 'Raw data row counts')
write.csv(raw_tbl, file.path(RESULT_PATH, 'raw_data_row_counts.csv'), row.names = FALSE)
```



```{r batch-plots}
# TODO: move plotting to end, plot PASS/FAIL as in qcpipeline2.R
# Batch plots - update to passed/failed
generic_plot_all_raw(
  raw.list = rawList,
  path = RESULT_PATH,
  write_pdf = TRUE,
  pdf_name = paste(RESULTS_NAME,"_all_dxf.pdf",sep="")
)
cat('Batch plot PDF written to:', file.path(RESULT_PATH, paste(RESULTS_NAME,"_all_dxf.pdf",sep="")), '\n')
```



```{r qaqc-run}
# Run QCIRMS QA/QC
start_time <- Sys.time()
currFiltered <- QAQC_IRMS(
  unfilteredPath = DATA_PATH,
  expRef.df = expRef.df,
  checkIntStand = CHECK_INTERNAL_STANDARDS,
  internalStandID = c('L1', 'H1', 'LW'),
  # add params: internal standard delta18O standard deviations and accepted values
  # LAC
  dataName = RESULTS_NAME,
  maxPkNum = 18,
  expectedNonSampPks = 7,
  sdCrefIso.thresh = 0.1,
  sdOrefIso.thresh = 0.1,
  checkRelDiffIntensity = TRUE,
  refSamp_relDiffInt.thresh = 0.75,
  amplName = 'Ampl44',
  ref_relDiffInt.thresh = 0.1,
  sdCsampIso.thresh = 0.3,
  sdOsampIso.thresh = 0.2,
  outPath = RESULT_PATH,
  verbose = TRUE
)
elapsed <- Sys.time() - start_time
cat('QA/QC runtime:', elapsed, '\n')
```

# QA/QC outputs

```{r qaqc-outputs}
refs.dat <- currFiltered[[1]]
samps.dat <- currFiltered[[2]]

cat('Reference rows after QA/QC:', nrow(refs.dat), '\n')
cat('Sample rows after QA/QC:', nrow(samps.dat), '\n')

knitr::kable(head(refs.dat), caption = 'Reference peaks passing QA/QC')
knitr::kable(head(samps.dat), caption = 'Sample peaks passing QA/QC')

# write.csv(refs.dat, file.path(RESULT_PATH, 'report_refs_passed.csv'), row.names = FALSE)
# write.csv(samps.dat, file.path(RESULT_PATH, 'report_samples_passed.csv'), row.names = FALSE)
if(length(currFiltered[[1]])!=0 & length(currFiltered[[2]]!=0)){
  summary_tbl <- data.frame(
    object = c('reference_peaks', 'sample_peaks'),
    n_rows = c(nrow(refs.dat), nrow(samps.dat)),
    n_cols = c(ncol(refs.dat), ncol(samps.dat)),
    stringsAsFactors = FALSE
  )
  knitr::kable(summary_tbl, caption = 'Summary of QA/QC outputs')
write.csv(summary_tbl, file.path(RESULT_PATH, 'report_summary.csv'), row.names = FALSE)
}

```

# Internal standards

```{r internal-standards}
if (CHECK_INTERNAL_STANDARDS && length(currFiltered) >= 3) {
  int_stand.list <- currFiltered[[3]]

  if (length(int_stand.list) >= 2 && !is.null(int_stand.list[[2]])) {
    knitr::kable(int_stand.list[[2]], caption = 'Average internal-standard deltas')
    write.csv(int_stand.list[[2]], file.path(RESULT_PATH, 'internal_standard_means.csv'), row.names = FALSE)
  }

  if (length(int_stand.list) >= 3 && !is.null(int_stand.list[[3]])) {
    knitr::kable(int_stand.list[[3]], caption = 'Internal-standard standard deviations')
    write.csv(int_stand.list[[3]], file.path(RESULT_PATH, 'internal_standard_sds.csv'), row.names = FALSE)
  }
} else {
  cat('Internal standards were not requested or were not returned.\n')
}
```

# visualize pass/fail in pdfs of chromatogram thumbnails
```{r pass-fail-pdfs}
# Get all filenames of .dxf files in the directory
passFiles <- unique(samps.dat$fileId)
if (!is.null(passFiles)){
  # get passed and failed files from previously defined file names
  passedInd <- which(fileNames %in% passFiles)
  failedFiles <- fileNames[-passedInd]

  # get intensity vs time data for passed and failed experiments
  passed_raw.list <- raw_data_all(passFiles, path=DATA_PATH)
  failed_raw.list <- raw_data_all(failedFiles, path=DATA_PATH)

  # plot all raw data and write to pdfs
  # passed QC
  generic_plot_all_raw(raw.list=passed_raw.list, path=PLOTS_PATH,
                     write_pdf=T,
                     pdf_name="passed_qc_spectra.pdf") 
  # failed QC
  generic_plot_all_raw(failed_raw.list, path=PLOTS_PATH,
                     write_pdf=T,
                     pdf_name="failed_qc_spectra.pdf") 
} else{
  cat("All files failed.\n")
}
```
